### 2020-11-23 written/V1_model/V1_model.tex
checkpoint V1 model

% You should title the file with a .tex extension (hw1.tex, for example)
\documentclass[11pt]{article}

\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{wrapfig}
\usepackage{fancyhdr}
\usepackage{tikz-qtree}
\usepackage{tikz-qtree-compat}
\usepackage[normalem]{ulem}
\usepackage{tikz}
\usepackage{systeme}
\usepackage{graphicx}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}

\oddsidemargin0cm
\topmargin-2cm     %I recommend adding these three lines to increase the 
\textwidth16.5cm   %amount of usable space on the page (and save trees)
\textheight23.5cm

\newcommand{\question}[2] {\vspace{.25in} \hrule\vspace{0.5em}
\noindent{\bf #1: #2} \vspace{0.5em}
\hrule \vspace{.10in}}
\renewcommand{\part}[1] {\vspace{.10in} {\bf (#1)}}

\newcommand{\myname}{Sean Bittner}
\newcommand{\myandrew}{srb2201@columbia.edu}
\newcommand{\myhwnum}{12}

\setlength{\parindent}{0pt}
\setlength{\parskip}{5pt plus 1pt}

\DeclarePairedDelimiter\abs{\lvert}{\rvert}%
 %
\pagestyle{fancyplain}
\rhead{\fancyplain{}{\myname\\ \myandrew}}

\begin{document}

\medskip                        % Skip a "medium" amount of space
                                % (latex determines what medium is)
                                % Also try: \bigskip, \littleskip

\thispagestyle{plain}
\begin{center}                  % Center the following lines
{\Large Draft of new V1 section} \\
Sean Bittner, Agostina Palmigiano \\
October 6, 2020 \\
\end{center}

\section{Exploratory analysis of V1 with EPI produces a novel theory}\label{results_V1}
Dynamical models of excitatory (E) and inhibitory (I) populations with supralinear input-output function have succeeded in explaining a host of experimentally documented phenomena.
In a regime characterized by inhibitory stabilization of strong recurrent excitation, these models give rise to paradoxical responses \cite{tsodyks1997paradoxical}, selective amplification  \cite{goldman2009memory, murphy2009balanced}, surround suppression \cite{ozeki2009inhibitory} and normalization \cite{rubin2015stabilized}. 
Despite their strong predictive power, E-I circuit models rely on the assumption that inhibition can be studied as an indivisible unit. 
However, experimental evidence shows that inhibition is composed of distinct elements -- parvalbumin (P), somatostatin (S), VIP (V) --
composing 80\% of GABAergic interneurons in V1 \cite{markram2004interneurons, rudy2011three, tremblay2016}, and that these inhibitory cell types follow specific connectivity patterns (Fig. \ref{fig:SV_flip}A) \cite{pfeffer2013inhibition}.
Recent theoretical advances \cite{litwin2016inhibitory, GarciaDelMolino2017, Chen2019},  have only started to address the consequences of this multiplicity in the dynamics of V1, strongly relying on linear theoretical tools. 
Here, we use EPI to gain a comprehensive understanding of how external circuit inputs govern S-V flipping and modulate variability.

We consider contrast responses of a nonlinear dynamical V1 circuit model (Fig. \ref{fig:SV_flip}A) with a state comprised of each neuron-type population's average membrane potential $x = \left[x_E, x_P , x_S, x_V \right]^\top$ (mV).
Specifically, we examine the stochastic stabilized supralinear network \cite{hennequin2018dynamical}, generalized to have inhibitory multiplicity (see Section \ref{methods_V1}).
Each population of neuron-type $\alpha$ receives recurrent input $\left(W f_r(x)\right)_\alpha$ from synaptic projections, where $W$ is the connectivity matrix and $f_r = \left[\cdot\right]_+^2$ is the rate nonlinearity.  
$W$ is calculated from recent experiments measuring post-synaptic potentials and connectivity rates in mice  \cite{allen2018layer, billeh2019systematic}.
Additionally, each population experiences a positive baseline input of $b_\alpha$, a contrast ($c$)-dependent input $h_\alpha$, and a stochastic input $\epsilon_\alpha$.
%Since contrast is only considered to change input to the E- and P-populations in V1 (\textbf{TODO} AP to add citation), $h_E = h_P = c h^{(c)}$ and $h_S = h_V = 0\text{mV}$.
Little is known about the scales of these baseline and contrast-dependent inputs, so we treat these as free parameters in our application of EPI:
 \begin{equation}
 z =  \begin{bmatrix} b_E & b_P & b_S & b_V & h^{(c)} \end{bmatrix}^\top.
\end{equation}

From first principles, we may attempt to fit or infer $z$ based on recorded V1 population responses at varying contrasts (\textbf{TODO} AP to add citation).
However, such a class of constrained models rarely produce good fits that generalize well.
Instead, we can use EPI to condition $z$ on one of the most salient properties of these data-sets called ``S-V flipping" --  a winner-take-all relationship between the S- and V-populations. At low contrast values, the V-population dominates the S-population, and at greater contrast values, the S-population dominates the V-population.  We formulate the emergent property of S-V flipping with a normalized statistic measuring the difference in steady state ($x_{ss}$) between the S- and V-population at a given contrast:
\begin{equation}
d_{S,V}(c; z) = \frac{f_r(x_{ss,S}(c; z)) - f_r(x_{ss,V}(c; z))}{|f_r(x_{ss}(c; z))|_2}
\end{equation}

\clearpage
\begin{figure}[h]
\caption{\small \textbf{A}.  Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons.   Some neuron-types largely do not form synaptic projections to others  (excitatory and inhibitory projections filled and unfilled, respectively).
\textbf{B}. EPI posterior $q_\theta(z \mid \mathcal{B}_{S-V})$ for S-V flipping. The obtained posterior is visualized as 500 samples from the inferred distribution colored by $\log(q_\theta(z))$. This posterior is bimodal and concentrated in planes $h^{(c)} > 0$ and $h^{(c)} < 0$ at distinct modes $z_1$ (black star) and $z_2$ (gray, star), respectively.  Bottom-left: Posterior predictive distribution of the emergent property statistics with respect to the constrained means (black, dashed line) and variances (gray, dashed lines at two standard deviations). 
\textbf{C}. Posterior predictive distribution of $d_{S,V}$ for each mode.  The $z_1$-mode produces V-to-S flipping with increasing contrast.
\textbf{D}. Model simulations at the mode $z_1$ at $c=0$ (solid) and $c=1$ (dashed). \textbf{E}. Mean system response at varying contrasts for $z_1$. Error bars show noise in rate. \textbf{F}. Plot of noise with contrast from E. Error bars are standard error measured across simulations.
 }\label{fig:SV_flip}
\begin{center}
\includegraphics[scale=.31]{figs/Fig3/Fig3.pdf}
\end{center}
\end{figure}
\clearpage

For S and V to flip their steady states, the difference between the S- and V-population rates at no contrast ($c=0$) must have opposite sign at full contrast ($c=1$).  Therefore, we stipulate the emergent property of S-V flipping to require the product of $d_{S,V}(c=0)$ and $d_{S,V}(c=1)$ to be appreciably negative.  The means and variance of this emergent property statistics were chosen such that the posterior predictive distribution always produced S-V flipping.

\begin{equation}
\begin{split}
\mathcal{B}_{S,V} \triangleq \text{\hspace{.5cm}} \mathbb{E}\left[ d_{S,V}(0)d_{S,V}(1) \right] &= -0.25 \\ 
\text{Var}\left[ d_{S,V}(0)d_{S,V}(1)\right] &=  0.125^2
\end{split}
\end{equation}

We ran EPI to obtain the parameter distribution of neuron-type population input parameters $z$ that produce S-V flipping (Fig. \ref{fig:SV_flip}B).
We considered positive baseline inputs $b_\alpha$ from 0 to 25mV and contrast-dependent changes in $h^{(c)}$ from -10 to 10mV.
The shape and structure of this distribution reveal some important properties of this model.
\begin{enumerate}
\item Two types of $z$ result in S-V flipping. (There are two modes separated by the hyperplane $h^{(c)} = 0$.)
\item S-V flipping is sensitive to anticorrelated changes in $b_S$ and $b_V$.  (The pairwise marginal distribution of $q_\theta(b_S, b_V \mid \mathcal{B}_{S,V})$ reveals strong correlation between these parameters.)
\item S-V flipping is most degenerate with respect to input to the P population. (The marginal distributions of $q_\theta(h_P \mid \mathcal{B}_{S,V})$ and  $q_\theta(dh_P \mid \mathcal{B}_{S,V})$ are approximately uniform along their allowed range.)
\end{enumerate}
Simulation from the posterior reveals that the biologically plausible mode ($h^{(c)} > 0$) matches experiments by producing V-to-S flipping with contrast.
Alternatively, the biologically implausible mode ($h^{(c)} > 0$) produces S-to-V flipping with contrast -- the opposite effect from experiments (Fig. \ref{fig:SV_flip}C).

We examined the responses of this circuit model to varying contrasts using the highest log-probability parameterization ($z_1$) of the plausible mode.
Simulations show the stochastic responses of the S- and V-population are indeed flipped at no and full contrast (Fig. \ref{fig:SV_flip}D).
This flipping occurs gradually with graded increases in contrast (Fig. \ref{fig:SV_flip}E).
Interestingly, the responses are noisier at intermediate contrasts (Fig. \ref{fig:SV_flip}F).
This suggests a new hypothesis explaining phenomena of variability in V1: noise is greater in stabilized supralinear networks during winner-take-all competition between the S- and V- populations, but is quenched as either the S- or V-population overtakes the other.

\section{V1 Supplemental section}\label{methods_V1}

The dynamics are driven by the rectified and exponentiated sum of recurrent inputs $Wx$, external inputs $h$, and external noise $\epsilon \sim \mathcal{N}(0, \sigma_{\epsilon}^2)$:

\begin{equation}
T \frac{dx}{dt} = -x + x_{\text{rest}} +Wf_r(x) + b + h + \epsilon
\end{equation}

where $T$ is a diagonal matrix with $\tau_E = 20\text{ms}$ and $\tau_P = \tau_S = \tau_V = 10\text{ms}$, $x_{\text{rest}} = -65\text{mV}$, and $f_r(x) = \left[x - x_{\text{rest}} \right]_+^2$.

The noise is an modeled as an Ornstein-Uhlenbeck process:
\begin{equation}
\tau_{\text{noise}} d\epsilon_\alpha = -\epsilon_\alpha dt + \sqrt{2\tau_{\text{noise}}}\sigma_\alpha dB
\end{equation}

where $\tau_{\text{noise}}$ is slow at 50ms, $\sigma_E = 0.2\text{mV}$, $\sigma_P = \sigma_S = \sigma_V = 0.1\text{mV}$, and $dB$ is a standard Wiener process.

We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1 \cite{allen2018layer, billeh2019systematic} (see Section \ref{methods_V1}).  The input to this circuit is comprised of a nominal baseline input $b$ and additive contrast-dependent input $h$
\begin{equation}
h = c \begin{bmatrix} h^{(c)} \\ h^{(c)} \\ 0 \\ 0 \end{bmatrix}.
\end{equation}

We considered fixed effective connectivity weights $W$ approximated from experimental recordings of publicly available datasets of mouse V1.  Specifically, Billeh et al. \cite{billeh2019systematic} produce estimates of the synaptic strength (in mV)
\begin{equation}
M = \begin{bmatrix} 
0.36 & -0.48 & -0.31 & -0.28 \\
1.49 & -0.68 & -0.50 & -0.18 \\
0.86 & -0.42 & -0.15 & -0.32 \\
1.31 & -0.41 & -0.52 & -0.37 \end{bmatrix}
\end{equation}
and connection probability
\begin{equation}
C = \begin{bmatrix} 0.16 & 0.411 & 0.424 &  0.087 \\
0.395 & .451 & 0.857 & 0.02 \\
0.182 & 0.03 & 0.082 & 0.625 \\
0.105 & 0.22 & 0.77 & 0.028 \end{bmatrix}.
\end{equation}

Multiplying these connection probabilities and synaptic efficacies gives us an effective connectivity matrix:
\begin{equation}
W = C \odot M = \begin{bmatrix} 
0.0576 & -0.197 & -0.131 & -0.0244 \\
0.589 & -0.307 & -0.429 & -0.00360 \\
0.157 & -0.0126  & -0.0123 & -0.200 \\
0.138 & -0.0902 & -0.400 & -0.0104 \end{bmatrix}.
\end{equation}

%\begin{figure}[h]
%\caption{\small \textbf{A}. ...
% }\label{fig:drdh}
%\begin{center}
%\includegraphics[scale=.3]{figs/FigSabc/FigSabc.pdf}
%\end{center}
%\end{figure}
%
%s
%\begin{figure}[h]
%\caption{\small \textbf{A}. ...
% }\label{fig:drdh}
%\begin{center}
%\includegraphics[scale=.15]{figs/FigSX/FigSX.pdf}
%\end{center}
%\end{figure}

\bibliography{epi}
\bibliographystyle{unsrt}

\end{document}

% With this model, we are interested in the differential responses of each neuron-type population to changes in input $dh$. 
% Initially, we studied the linearized response of the system to input $\frac{dx_{ss}}{dh}$ at the steady state response $x_{ss}$, i.e. a fixed point. 
% All analyses of this model consider the steady state response, so we drop the notation $ss$ from here on.
% While this linearization accurately predicts differential responses $dx = \left[ dx_{E} , dx_{P} , dx_{S} ,dx_{V} \right]^\top$  for small differential inputs to each population $dh = \left[ 0.1 , 0.1 , 0.1 , 0.1 \right]^\top$ (Fig \ref{fig:V1_EPI}B left), the linearization is a poor predictor in this nonlinear model more generally (Fig. \ref{fig:V1_EPI}B right).  
% Currently available approaches to deriving the steady state response of the system are limited.

%\section{Stability analysis}
%The input $h = b + dh$ is comprised of a baseline input  $b = \left[ b_E, b_P , b_S , b_V \right]^\top = \left[ 1, 1, 1 ,1.25 \right]^\top$  and a differential input $dh = \left[ dh_E , dh_P , dh_S , dh_V\right]^\top$ to each neuron-type population.  
%
%We want to know the differential inputs $dh$ that maintain the steady state $x_{\alpha}$ for $\alpha \in \{E, P, S, V\}$. 
%We see from Figure 1B that input to a single population in the recurrent circuit elicits a variety of responses across populations: E same, P up, S down, and V up.
%We define the differential steady state $dx_{\alpha}$ as the change in steady state $x_{\alpha}$ when receiving input $h=b + dh$ with respect to the baseline $h = b$.
%Maintaining the steady state of a neuron-type population amounts to the emergent property 
%\begin{equation}
%\mathcal{B}(\alpha, \sigma) ~~\triangleq~~ 
%\mathbb{E} \begin{bmatrix} dx_{\alpha} \\ dx_{\alpha}^2 \end{bmatrix} ~~=~~ \begin{bmatrix} 0 \\ \sigma^2 \end{bmatrix}.\\
%\end{equation}
%In the following analyses, we chose $\sigma=0.25$.
%
%\section{EPI agrees with ABC}
%To get an idea of what distribution of parameters ($dh$) we should expect from EPI, we can use ABC to obtain a set of parameters related to the emergent property.  
%We compare EPI to ABC with a rejection heuristic defined by the standard deviation of the differential responses $\sigma_{ABC}$
% \[f_{ABC}(dx_\alpha; \sigma_{ABC}) = |dx_\alpha| > 2\sigma_{ABC}.\]
% In other words, we ran ABC accepting parameters that generate differential responses within two standard deviations $\sigma_{ABC}=0.25$ of $dx_\alpha = 0$.
% In Figure 2, we see that the distributions obtained via EPI (colored by $\log q_\theta(z)$) are visually similar to those obtained via ABC (colored by $dx_\alpha$). \\
% 
% \textbf{Figure 2}: EPI (left) vs ABC (right).  Arrows in EPI distribution indicate dimensions of maximal sensitivity at selected parameters $dh$. The importance of $v_S$ and $v_P$ are explained in section 4.2. \\
%
%There are some subtle differences between the EPI and ABC distributions, but some difference is to be expected.
%Rather than simply attributing these differences to imperfection of the EPI optimization routine, we consider the the effect of bias in $dx_\alpha$ from ABC (and lack thereof from EPI).
%
%When $\epsilon > 0$ in ABC, the samples are \textit{not} from the posterior distribution.
%For $\epsilon > 0$, the ``posterior" predictive means of ABC $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha \right]$ may be far from zero.
%In Figure 3, we see that with ABC, there is an increasingly negative bias in $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha \right]$ for greater error tolerances across all neuron-types.  
%Additionally with ABC, there is no precise control of the variance $\mathbb{E}_{ABC(\alpha, \sigma_{ABC})}\left[dx_\alpha^2 \right]$, which may be undesireable.
%
%In contrast, the first and second moments of $dx_\alpha$ are controlled to a specified degree of accuracy (Figure 4).
%The variances of each distribution of $dx_\alpha$ are close to their target value $0.25^2 = 0.0625$ (see variances next to histograms of Figure 4).